---
title: "FinalLehmann"
author: "Ben Lehmann"
date: "2023-05-08"
output: html_document
---
https://github.com/BenLehmann12/Student-Performance
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We have a dataset of test scores (Math, Reading, Writing), with different variables, Gender, Race, If they have college education or not, if they have completed test prep.

Now I want to look at the average score (Math+Reading+Writing/3) and compare them by race and gender

I also want to look at Scores by Gender
There is not much data to clean.
-Uses: Histogram, Bar Graph, Boxplot, Scatterplot

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
```

```{r}
students <- read_csv("~/DS202RWork/StudentsPerformance.csv")


namesOfColumns <- c("Gender","Race","Parent_Education","Lunch","Test_Prep","Math_Score","Reading_Score","Writing_Score")
colnames(students) <- namesOfColumns

students <- students %>% mutate (Avg.score = (Reading_Score + Math_Score + Writing_Score)/3)
```
```{r}
students %>% group_by(Gender)%>% summarise(n=n()) %>% ggplot(aes(Gender,n,fill=Gender))+geom_col()+geom_text(aes(label = n))
```
```{r}
students %>% group_by(Race, Gender) %>% summarise(n = n()) %>% 
  ggplot(aes(x = Race, y = n, fill = Gender)) + 
  geom_col(position = 'dodge')
```
The Graphs above are general population, I just want to see gender and race together, men and women by race
Now to get a good perspective on scores by gender, boxplots are great for side by side analysis

```{r}
g1 <- ggplot(students, aes(Gender, Math_Score)) + geom_boxplot() +xlab("Gender") + ylab("Math Scores")
g1
```
In terms of Math, Males have better Scores than Women, With little outliers and higher median

```{r}
g2 <- ggplot(students, aes(Gender, Reading_Score)) + geom_boxplot() +xlab("Gender") + ylab("Reading Scores")
g2
```
In Terms of Reading Scores, Women tend to have higher Median, Average than Men

```{r}
g3 <- ggplot(students, aes(Gender, Writing_Score)) + geom_boxplot() +xlab("Gender") + ylab("Writing Scores")
g3
```
Based on the Boxplot Above, Women have higher scores than Men, Men have a much lower Median than Women and Much lower Average.

As we have seen the boxplots, we can see the distribution of scores for Men and Women and we can see Men have an edge in higher scores. Let us look at them individually. 

```{r}
newP <-ggplot(students, aes(Math_Score)) + geom_histogram(binwidth=5) + facet_wrap(~Gender)
```
```{r}
wP <-ggplot(students, aes(Writing_Score)) + geom_histogram(binwidth=5) + facet_wrap(~Gender)
wP
```

```{r}
rP <-ggplot(students, aes(Reading_Score)) + geom_histogram(binwidth=5) + facet_wrap(~Gender)
rP
```


Now look at them together.
```{r}
p <- ggplot(students, aes(Math_Score)) + geom_histogram(binwidth=5, color="black", aes(fill=Gender))
p
```

```{r}
p_new <- ggplot(students, aes(Reading_Score)) + geom_histogram(binwidth=5, color="black", aes(fill=Gender))
p_new
```

```{r}
p3 <- ggplot(students, aes(Writing_Score)) + geom_histogram(binwidth=5, color="black", aes(fill=Gender))
p3
```

```{r}
students %>% group_by(Race) %>% summarise(new_mean=round(sum(Avg.score)/n(),2)) %>% ggplot(aes(Race,new_mean,fill=Race))+geom_bar(stat='identity')+geom_text(aes(label = new_mean)) +labs("Mean Test Score")
```

Now, Let's look at the average score by Gender, We can see the Females perform better than Men in Testing

```{r}
students %>% group_by(Gender) %>% summarise(new_mean=round(sum(Avg.score)/n(),2)) %>% ggplot(aes(Gender,new_mean,fill=Gender))+geom_bar(stat='identity')+geom_text(aes(label = new_mean))
```
When we look at the Graphs above, we can see that Women have a better Average Score compared to Men. And that Group E(White) have a higher Averge Test Score compared to other minority groups.
Then to find only Male data and Female data, we use filter()

```{r}
girl <- students %>% filter(Gender=="female")
boy <- students %>% filter(Gender=="male")

```


```{r}
ggplot()+geom_point(girl,mapping=aes(Math_Score,Reading_Score,color='female'))+geom_point(boy,mapping=aes(Math_Score,Reading_Score,color='male'))
```

```{r}
ggplot()+geom_point(girl,mapping=aes(Writing_Score,Reading_Score,color='female'))+geom_point(boy,mapping=aes(Writing_Score,Reading_Score,color='male'))
```




Let's look at a Correlation between Math Score, Men and Women with Good Math Scores tend to have Good Reading Score, and vice versa.

Conclusion:
When looking at each type of scores indivdually by Gender, We see that Men do better than the Women based on the boxplots, however, Women do better than Men in reading and writing when looking at the boxplots. Once we average the scores, women have the upper hand than Men, (69.57 to 65.84).
